Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2023-01-06 supplement.tex
final single origin best fit values

\texttt{moments-LD} uses a composite multivariate Gaussian likelihood approach 
to simultaneously fit relative pairwise diversity and LD statistics over a range of 
recombination distances.
This composite likelihood and Gaussian assumption are described in more detail 
 and validated in \citet{Ragsdale2019-nt}, but we outline them here.  
To compute likelihoods, we need an estimate of the joint
distribution of summary statistics. 
We estimate uncertainty due to the finite amount of genetic material
used in inference using bootstrap over 500 segments along the genome with
roughly equal lengths of retained sequences within each segment. First, for
each distance bin, we use these bootstrap samples to obtain a
variance-covariances matrix across all statistics. This variance-covariance
matrix is used to obtain a model likelihood for each recombination distance
bin and single-locus nucleotide diversity, as a multivariate Gaussian
likelihood. 
The full model likelihood is taken as the product of likelihoods
over each bin. In other words, we followed the approach of  \citet{Ragsdale2019-nt} 
in optimizing a composite likelihood where
observations in different bins were taken to be independent.

\aprcomment{These paragraphs are outdated.. we drop Godambe. need to fix}

For the best fit parameters, we computed $95\%$ confidence intervals using the
same set of bootstrapped samples and Godambe Information approach
\citep{Coffman2016-yq}, which corrects composite likelihood estimates of
confidence intervals to account for nonindependence in the data, including
linkage between loci and nonindependence of recombination bins.  To ensure that
our confidence intervals computed using the approach from
\citet{Coffman2016-yq} do not underestimate true CIs, we also performed block
bootstrap resampling of our data, 200 times, and refit each model to the
bootstrapped datasets. From these, we evaluated the distribution of inferred
parameters, which should be similar to or narrower than the CIs estimated from
the GIM approach (Section~\ref{sec:bootstrap-CIs}).

Tables~\ref{tab:supp-single-origin}--\ref{tab:supp-merger-with-stem-migration}
give the best fit values for all parameters in the four models considered here,
along with standard errors using computed using the Godambe method.
Figures~\ref{fig:supp-single-origin-fits}--\ref{fig:supp-merger-with-stem-migration-fits}
show the corresponding fits to many of the LD statistics used in the inference.
The $\sigma_{Dz}$ statistics are fit poorly by the single-origin model, while
the models with multiple stems all provide reasonable fits to the data. The
merger-with-stem-migration model performed the best of these four models,
although the fits were nearly visually indistinguishable from the fits from the
merger-with-stem-migration model.

@@

\begin{figure}[htb!]
    \centering
    \includegraphics{figures/supp-merged-data-model-fits}
    \caption{
        \textbf{Merged datasets bias demographic inferences.}
        The panels represent best-fit models according to 
        different combinations of high- and low- coverage data.
        While statistics from low and high coverage datasets provide consistent
        best-fit demographic models, merged datasets result in biased LD
        statistics (Figure~\ref{fig:supp-merged-data-SEs}). These biased LD
        statistics lead to large downstream biases in inferred demographic
        models. 
    }
    \label{fig:supp-merged-data-fits}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    \includegraphics{./figures/supp-refit-lls.pdf}
    \caption{
        \textbf{Log-likelihoods of fits of four models to simulated data.} In
        three scenarios, a simple single origin model and our preferred
        continuous migration and merger with stem migration models, we
        simulated 500 replicates of 500 Mb with the same populations and sample
        sizes from our reported best fit models. We then fit the four models
        reported in the main text to each of these simulated datasets and
        compared log-likelihoods. In each case, the simulated model is
        correctly chosen, even under the single origin scenario. We conclude
        our more complex models (with more parameters) are not simply fitting
        statistical noise.
    }
    \label{fig:refit-lls}
\end{figure}

\subsection{Evaluation of confidence intervals}
\label{sec:bootstrap-CIs}

\aprcomment{This section will be removed}

We created 200 block-bootstrap replicate datasets and refit our four reported
models to each replicate (Section~\ref{sec:likelihoods-uncertainty}). From
these refit models, we estimated the empirical uncertainty by taking the
standard deviation of the distribution of each refit parameter. We compared
these to our estimated standard errors from the Godambe Information Matrix
(GIM) approach to check that the confidence intervals from the GIM approach do
not underestimate true uncertainty (Table~\ref{tab:supp-confidence-intervals}).

The GIM-computed standard errors are generally close to those from the fits to
the bootstrap data, and some are larger (roughly 2--5x) than those from the
bootstrap fits. A possible explanation for this discrepancy is that some
parameters of our models come close to the lower and upper bounds allowed for
those parameters (e.g., small recent population size in the Nama, a bottleneck
in one of the stems). In such cases, the distribution of
parameters under constraints, and that of other correlated parameters in the model
may be narrower than
the true (unconstrained) uncertainties in those parameters. 
%Model constraints and parameter
%bounds in optimization may also lead to parameters fits that are away from
%their true optimal value. In this case, the GIM-estimated confidence intervals
%should be larger and contain what would be found from an unconstrained fit.

To demonstrate this effect, we simulated data under a simple
isolation-with-migration model, in which population sizes are allowed to vary,
and one population has a severe bottleneck in the most recent 1000 years
(Figure~\ref{fig:supp-uncerts-toy}). We simulated 200 1Mb regions with
$u=r=10^{-8}$, sampling 10 diploid individuals from each present-day population
and computed LD statistics. From these, we inferred the best fit parameters
without imposing bounds on the parameters, and we also reinferred the model
imposing a lower bound of 300 on population sizes (the bottleneck size). From
our fits, we computed confidence intervals using the Fisher information matrix
(FIM) and GIM approaches, in both the unbounded and bounded fits. We also
reinferred parameters over 200 bootstrapped replicate datasets to compare
empirical uncertainties to those estimated CIs.

As expected the FIM CIs underestimate the true uncertainty and the GIM CIs
closely match empirical uncertainty from the bootstrap fits. In the fit that
imposed a lower bound on population sizes, some GIM-estimated standard errors
were larger than those from the unbounded fits.  However, the fit parameters
were slightly shifted from the unbounded fits, due to hitting the lower bound
of the bottleneck size. In these cases, the estimated CIs were larger but
contained the true CIs from the unbounded GIM and bootstrap fits. For parameters
that were not strongly affected by hitting the bounds, the CIs matched closely
(Figure~\ref{fig:supp-uncerts-toy}).

\begin{figure}[htb]
    \centering
    \includegraphics{./figures/supp-uncerts-toy.pdf}
    \caption{
        \textbf{Evaluation of GIM confidence intervals using moments-LD.}
        The GIM approach accurately estimates confidence intervals, accounting
        for non-independence between regions of linked loci. GIM-estimated CIs
        from the constrained fit are large enough to contain the true
        uncertainty. Vertical dashed lines indicate simulated parameter values.
    }
    \label{fig:supp-uncerts-toy}
\end{figure}

@@

\begin{table}[ht]
\caption{
    \label{tab:supp-single-origin}
    \textbf{Best-fit parameters from the single-origin model.}
    This model includes a single stem population that splits into contemporary human
    groups. Fixed parameters are specified in Table~\ref{tab:fixed-params}.
    Inferred values are scaled to physical units assuming a generation time of
    29 years. This model gave a log-likelihood of -189,434.
}
\centering
\begin{tabular}[t]{rP{8cm}cc}
    \toprule
    Parameter & Description & Value & 95\% CI \\
    \midrule
    $N_e$ & Ancestral effective population size & 9910 & 9600--10090 \\
    $N_{MH}$ & Size of human lineage between Neanderthal and Nama splits & 20890 & 20640--21220 \\
    $N_{Nama_0}$ & Initial Nama size & 10170 & 9950--10250 \\
    $N_{Nama_F}$ & Final Nama size & 223 & 217--226 \\
    $N_{MSL_0}$ & Initial Mende size & 17250 & 16770--17490 \\
    $N_{MSL_F}$ & Final Mende size & 16760 & 16370--17060 \\
    $N_{EA}$ & Size of East African branch & 7130 & 6940--7130 \\
    $N_{Gumuz_F}$ & Final Gumuz size & 3830 & 3690--3900 \\
    $N_{EP}$ & East African agriculturist size & 13010 & 12720--13140 \\
    $N_{GBR_0}$ & Initial British size & 843 & 836--862 \\
    $N_{GBR_F}$ & Final British size & 12150 & 11870--12300 \\
    $N_{Neand}$ & Neanderthal size & 1850 & 1650--1990 \\
    $T_{Nama}$ & Nama split time (years) & 110700 & 107800--112200 \\
    $m_{Nama-MSL}$ & Nama--Mende symmetric migration rate & $2.82\times10^{-5}$ & $2.61\times10^{-5}$--$3.02\times10^{-5}$ \\
    $m_{Nama-EA}$ & Nama--East Africa symmetric migration rate & $4.98\times10^{-5}$ & $4.82\times10^{-5}$--$5.31\times10^{-5}$ \\
    $m_{MSL-EA}$ & Mende--East Africa migration rate & $18.7\times10^{-5}$ & $18.7\times10^{-5}$--$19.2\times10^{-5}$ \\
    $m_{EA-GBR}$ & East Africa--Europe migration rate & $4.42\times10^{-5}$ & $4.13\times10^{-5}$--$4.54\times10^{-5}$ \\
    $m_{EA-EA}$ & Intra-East Africa migration rate & $41.4\times10^{-5}$ & $41.0\times10^{-5}$--$44.2\times10^{-5}$ \\
    $f_{GBR \rightarrow EP}$ & Ancestry proportion of East African agriculturalists from GBR 12ka ($1-f$ from Gumuz) & 0.659 & 0.657--0.671 \\
    $f_{EP \rightarrow Nama}$ & Ancestry proportion from EA pastoralists to Nama 2ka & 0.280 & 0.270--0.286 \\
    $f_{GBR \rightarrow Nama}$ & Ancestry proportion from Europeans to Nama 10 generations ago & 0.150 & 0.147--0.154 \\
    \bottomrule
\end{tabular}
\end{table}

@@

\begin{table}[ht]
\caption{
    \label{tab:supp-merger-without-stem-migration}
    \textbf{Best-fit parameters from the Merger-Without-Stem-Migration model.}
    Inferred values are scaled to physical units assuming a generation time of
    29 years. This model gave a log-likelihood of -107,652.
}
\centering
\begin{tabular}[t]{rP{8cm}cc}
    \toprule
    Parameter & Description & Value & 95\% CIs\\
    \midrule
    $N_e$ & Ancestral effective population size & 11300 & 10100--11290 \\
    $N_{stem1}$ & Size of stem 1 lineage between stem 1--stem 2 split and stem 1E--stem 1S split & 113 & 113--606 \\
    $N_{stem2}$ & Size of stem 2 lineage & 24780 & 26500--35060 \\
    $N_{Nama_0}$ & Initial Nama and stem 1S size & 13080 & 11310--12910 \\
    $N_{Nama_F}$ & Final Nama size & 224 & 223--238 \\
    $N_{MSL_0}$ & Initial Mende size & 11870 & 10280--11630 \\
    $N_{MSL_F}$ & Final Mende size & 25570 & 25720--30180 \\
    $N_{EA}$ & Size of East African and stem 1E branch & 9110 & 8680--9140 \\
    $N_{Gumuz_F}$ & Final Gumuz size & 3370 & 3270--3470 \\
    $N_{EP}$ & East African agriculturist size & 13550 & 13120--13690 \\
    $N_{GBR_0}$ & Initial British size & 927 & 920--963 \\
    $N_{GBR_F}$ & Final British size & 12050 & 11790--12430 \\
    $N_{Neand}$ & Neanderthal size & 1940 & 1770--2110 \\
    $T_{stems}$ & Stem split time (years) & 411700 & 381400--410700 \\
    $T_{stem1}$ & Stem 1 split time into stem 1E and stem 1S (years) & 355400 & 154300--320400 \\
    $m_{Nama-MSL}$ & Nama--Mende symmetric migration rate & $0.361\times10^{-5}$ & $0.205\times10^{-5}$--$0.697\times10^{-5}$ \\
    $m_{Nama-EA}$ & Nama--East Africa symmetric migration rate & $4.04\times10^{-5}$ & $3.90\times10^{-5}$--$4.42\times10^{-5}$ \\
    $m_{MSL-EA}$ & Mende--East Africa migration rate & $19.3\times10^{-5}$ & $19.8\times10^{-5}$--$20.8\times10^{-5}$ \\
    $m_{EA-GBR}$ & East Africa--Europe migration rate & $3.76\times10^{-5}$ & $3.34\times10^{-5}$--$3.95\times10^{-5}$ \\
    $m_{EA-EA}$ & Intra-East Africa migration rate & $37.2\times10^{-5}$ & $35.2\times10^{-5}$--$39.6\times10^{-5}$ \\
    $f_{GBR \rightarrow EP}$ & Ancestry proportion of East African agriculturalists from GBR 12ka ($1-f$ from Gumuz) & 0.648 & 0.641--0.661 \\
    $f_{EP \rightarrow Nama}$ & Ancestry proportion from EA pastoralists to Nama 2ka & 0.258 & 0.245--0.264 \\
    $f_{GBR \rightarrow Nama}$ & Ancestry proportion from Europeans to Nama 10 generations ago & 0.156 & 0.152--0.161 \\
    $m_{stems}$ & Stem 1--stem 2 migration rate & $0$ & fixed \\
    $T_{Nama}$ & Time of Nama merger event & 115400 & 101400--116800 \\
    $f_{stem 2 \rightarrow Nama}$ & Proportion of stem 2 ancestry making up initial Nama lineage ($1-f$ from stem 1S) & 0.693 & 0.706--0.769 \\
    $T_{EA}$ & Time of East Africa merger event & 92940 & 94500--10140 \\ %%% outside CI!
    $f_{stem 2 \rightarrow EA}$ & Proportion of stem 2 ancestry making up initial East Africa lineage ($1-f$ from stem 1E) & 0.502 & 0.510--0.564 \\
    $T_{MSL}$ & Time of secondary admixture from stem 2 to Mende & 23490 & 22920--25510 \\
    $f_{stem 2 \rightarrow MSL}$ & Proportion of ancestry from secondary stem 2 admixture to Mende & 0.170 & 0.174--0.196 \\
    \bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{
    \label{tab:supp-merger-with-stem-migration}
    \textbf{Best-fit parameters from the Merger-With-Stem-Migration model.}
    Inferred values are scaled to physical units assuming a generation time of
    29 years. This model gave a log-likelihood of -102,633. \aprcomment{check these at end}
    \aprcomment{if the slightly shift, also fix elsewhere in supp and main text}
}
\centering
\begin{tabular}[t]{rP{8cm}cc}
    \toprule
    Parameter & Description & Value & Std. err.\\
    \midrule
    $N_e$ & Ancestral effective population size & 10960 & 64500--11910 \\
    $N_{stem1}$ & Size of stem 1 lineage between Neanderthal split and stem 1E--stem 1S split & 100 & 100--851 \\
    $N_{stem2}$ & Size of stem 2 lineage & 24100 & 22130--26340 \\
    $N_{Nama_0}$ & Initial Nama size & 13530 & 11500--13560 \\
    $N_{Nama_F}$ & Final Nama size & 219 & 214--231 \\
    $N_{MSL_0}$ & Initial Mende size & 11500 & 9690--11290 \\
    $N_{MSL_F}$ & Final Mende size & 27390 & 26820--32280 \\
    $N_{EA}$ & Size of East African Branch & 9150 & 7970--8860 \\
    $N_{Gumuz_F}$ & Final Gumuz size & 3390 & 3270--3570 \\
    $N_{EP}$ & East African agriculturist size & 13500 & 12780--13500 \\
    $N_{GBR_0}$ & Initial British size & 950 & 928--1002 \\
    $N_{GBR_F}$ & Final British size & 12370 & 11850--12600 \\
    $N_{Neand}$ & Neanderthal size & 2510 & 2100--2730 \\
    $T_{stems}$ & Stem split time (years) & 1475000 & 695000--1785000 \\
    $T_{stem1}$ & Stem 1S--stem 1E split time (years) & 479100 & 275600--478300 \\
    $m_{Nama-MSL}$ & Nama--Mende symmetric migration rate & $1.06\times10^{-5}$ & $0.66\times10^{-5}$--$1.57\times10^{-5}$ \\
    $m_{Nama-EA}$ & Nama--East Africa symmetric migration rate & $4.49\times10^{-5}$ & $4.45\times10^{-5}$--$5.52\times10^{-5}$ \\
    $m_{MSL-EA}$ & Mende--East Africa migration rate & $19.6\times10^{-5}$ & $20.0\times10^{-5}$--$21.7\times10^{-5}$ \\
    $m_{EA-GBR}$ & East Africa--Europe migration rate & $3.85\times10^{-5}$ & $3.25\times10^{-5}$--$4.27\times10^{-5}$ \\
    $m_{EA-EA}$ & Intra-East Africa migration rate & $36.0\times10^{-5}$ & $33.7\times10^{-5}$--$40.9\times10^{-5}$ \\
    $f_{GBR \rightarrow EP}$ & Ancestry proportion of East African agriculturalists from GBR 12ka ($1-f$ from Gumuz) & 0.639 & 0.636--0.667 \\
    $f_{EP \rightarrow Nama}$ & Ancestry proportion from EA pastoralists to Nama 2ka & 0.259 & 0.237--0.267 \\
    $f_{GBR \rightarrow Nama}$ & Ancestry proportion from Europeans to Nama 10 generations ago & 0.159 & 0.154--0.167 \\
    $m_{stems}$ & Stem 1--stem 2 migration rate & $11.8\times10^{-5}$ & $3.95\times10^{-5}$--$12.3\times10^{-5}$ \\
    $T_{Nama}$ & Time of Nama merger event & 121600 & 101300--124600 \\
    $f_{stem 2 \rightarrow Nama}$ & Proportion of stem 2 ancestry making up initial Nama lineage ($1-f$ from stem 1S) & 0.692 & 0.694--0.758 \\
    $T_{EA}$ & Time of East Africa merger event & 96500 & 95400--111000 \\
    $f_{stem 2 \rightarrow EA}$ & Proportion of stem 2 ancestry making up initial East Africa lineage ($1-f$ from stem 1E) & 0.511 & 0.514--0.589 \\
    $T_{MSL}$ & Time of secondary admixture from stem 2 to Mende & 24890 & 22210--26300 \\
    $f_{stem 2 \rightarrow MSL}$ & Proportion of ancestry from secondary stem 2 admixture to Mende & 0.182 & 0.182--0.211 \\
    \bottomrule
\end{tabular}
\end{table}

>>>>

\texttt{moments-LD} uses a composite multivariate Gaussian likelihood approach
to simultaneously fit relative pairwise diversity and LD statistics over a
range of recombination distances. This composite likelihood and Gaussian
assumption are described in more detail and validated in
\citet{Ragsdale2019-nt}, but we outline them here. To compute likelihoods, we
need an estimate of the joint distribution of summary statistics. We estimate
uncertainty due to the finite amount of genetic material used in inference
using bootstrap over 500 segments along the genome with roughly equal lengths
of retained sequences within each segment. First, for each distance bin, we use
these bootstrap samples to obtain a variance-covariances matrix across all
statistics. This variance-covariance matrix is used to obtain a model
likelihood for each recombination distance bin and single-locus nucleotide
diversity, as a multivariate Gaussian likelihood. The full model likelihood is
taken as the product of likelihoods over each bin. In other words, we followed
the approach of \citet{Ragsdale2019-nt} in optimizing a composite likelihood
where observations in different bins were taken to be independent. We computed
$95\%$ confidence intervals for each model using inferences from 200
block-bootstrapped datasets, which were fit using the same procedure as the
original data. From the resulting parameters from these fits, we removed the
bottom $2.5\%$ and top $97.5\%$ values and report the range of the remaining
fit parameter values.

Tables~\ref{tab:supp-single-origin}--\ref{tab:supp-merger-with-stem-migration}
give the best fit values for all parameters in the four models considered here,
along with 95\% confidence intervals from the refit bootstrapped datasets.
Figures~\ref{fig:supp-single-origin-fits}--\ref{fig:supp-merger-with-stem-migration-fits}
show the corresponding fits to many of the LD statistics used in the inference.
The $\sigma_{Dz}$ statistics are fit poorly by the single-origin model, while
the models with multiple stems all provide reasonable fits to the data. The
merger-with-stem-migration model performed the best of these four models,
although the fits were nearly visually indistinguishable from the fits from the
merger-with-stem-migration model.

@@

\begin{figure}[ht!]
    \centering
    \includegraphics{figures/supp-merged-data-model-fits}
    \caption{
        \textbf{Merged datasets bias demographic inferences.}
        The panels represent best-fit models according to 
        different combinations of high- and low- coverage data.
        While statistics from low and high coverage datasets provide consistent
        best-fit demographic models, merged datasets result in biased LD
        statistics (Figure~\ref{fig:supp-merged-data-SEs}). These biased LD
        statistics lead to large downstream biases in inferred demographic
        models. 
    }
    \label{fig:supp-merged-data-fits}
\end{figure}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics{./figures/supp-refit-lls.pdf}
    \caption{
        \textbf{Log-likelihoods of fits of four models to simulated data.} In
        three scenarios, a simple single origin model and our preferred
        continuous migration and merger with stem migration models, we
        simulated 500 replicates of 500 Mb with the same populations and sample
        sizes from our reported best fit models. We then fit the four models
        reported in the main text to each of these simulated datasets and
        compared log-likelihoods. In each case, the simulated model is
        correctly chosen, even under the single origin scenario. We conclude
        our more complex models (with more parameters) are not simply fitting
        statistical noise.
    }
    \label{fig:refit-lls}
\end{figure}

@@

\begin{table}[ht]
\caption{
    \label{tab:supp-single-origin}
    \textbf{Best-fit parameters from the single-origin model.}
    This model includes a single stem population that splits into contemporary human
    groups. Fixed parameters are specified in Table~\ref{tab:fixed-params}.
    Inferred values are scaled to physical units assuming a generation time of
    29 years. This model gave a log-likelihood of -189,434.
}
\centering
\begin{tabular}[t]{rP{8cm}cc}
    \toprule
    Parameter & Description & Value & 95\% CI \\
    \midrule
    $N_e$ & Ancestral effective population size & 9860 & 9600--10090 \\
    $N_{MH}$ & Size of human lineage between Neanderthal and Nama splits & 20850 & 20640--21220 \\
    $N_{Nama_0}$ & Initial Nama size & 10170 & 9950--10250 \\
    $N_{Nama_F}$ & Final Nama size & 222 & 217--226 \\
    $N_{MSL_0}$ & Initial Mende size & 17280 & 16770--17490 \\
    $N_{MSL_F}$ & Final Mende size & 16730 & 16370--17060 \\
    $N_{EA}$ & Size of East African branch & 7130 & 6940--7130 \\
    $N_{Gumuz_F}$ & Final Gumuz size & 3800 & 3690--3900 \\
    $N_{EP}$ & East African agriculturist size & 12990 & 12720--13140 \\
    $N_{GBR_0}$ & Initial British size & 845 & 836--862 \\
    $N_{GBR_F}$ & Final British size & 12140 & 11870--12300 \\
    $N_{Neand}$ & Neanderthal size & 1850 & 1650--1990 \\
    $T_{Nama}$ & Nama split time (years) & 110900 & 107800--112200 \\
    $m_{Nama-MSL}$ & Nama--Mende symmetric migration rate & $2.83\times10^{-5}$ & $2.61\times10^{-5}$--$3.02\times10^{-5}$ \\
    $m_{Nama-EA}$ & Nama--East Africa symmetric migration rate & $4.98\times10^{-5}$ & $4.82\times10^{-5}$--$5.31\times10^{-5}$ \\
    $m_{MSL-EA}$ & Mende--East Africa migration rate & $18.7\times10^{-5}$ & $18.7\times10^{-5}$--$19.2\times10^{-5}$ \\
    $m_{EA-GBR}$ & East Africa--Europe migration rate & $4.38\times10^{-5}$ & $4.13\times10^{-5}$--$4.54\times10^{-5}$ \\
    $m_{EA-EA}$ & Intra-East Africa migration rate & $43.8\times10^{-5}$ & $41.0\times10^{-5}$--$44.2\times10^{-5}$ \\
    $f_{GBR \rightarrow EP}$ & Ancestry proportion of East African agriculturalists from GBR 12ka ($1-f$ from Gumuz) & 0.661 & 0.657--0.671 \\
    $f_{EP \rightarrow Nama}$ & Ancestry proportion from EA pastoralists to Nama 2ka & 0.280 & 0.270--0.286 \\
    $f_{GBR \rightarrow Nama}$ & Ancestry proportion from Europeans to Nama 10 generations ago & 0.150 & 0.147--0.154 \\
    \bottomrule
\end{tabular}
\end{table}

@@

\begin{table}[ht]
\caption{
    \label{tab:supp-merger-without-stem-migration}
    \textbf{Best-fit parameters from the Merger-Without-Stem-Migration model.}
    Inferred values are scaled to physical units assuming a generation time of
    29 years. This model gave a log-likelihood of -107,652.
}
\centering
\begin{tabular}[t]{rP{8cm}cc}
    \toprule
    Parameter & Description & Value & 95\% CIs\\
    \midrule
    $N_e$ & Ancestral effective population size & 11300 & 10100--11290 \\
    $N_{stem1}$ & Size of stem 1 lineage between stem 1--stem 2 split and stem 1E--stem 1S split & 113 & 113--606 \\
    $N_{stem2}$ & Size of stem 2 lineage & 24780 & 26500--35060 \\
    $N_{Nama_0}$ & Initial Nama and stem 1S size & 13080 & 11310--12910 \\
    $N_{Nama_F}$ & Final Nama size & 224 & 223--238 \\
    $N_{MSL_0}$ & Initial Mende size & 11870 & 10280--11630 \\
    $N_{MSL_F}$ & Final Mende size & 25570 & 25720--30180 \\
    $N_{EA}$ & Size of East African and stem 1E branch & 9110 & 8680--9140 \\
    $N_{Gumuz_F}$ & Final Gumuz size & 3370 & 3270--3470 \\
    $N_{EP}$ & East African agriculturist size & 13550 & 13120--13690 \\
    $N_{GBR_0}$ & Initial British size & 927 & 920--963 \\
    $N_{GBR_F}$ & Final British size & 12050 & 11790--12430 \\
    $N_{Neand}$ & Neanderthal size & 1940 & 1770--2110 \\
    $T_{stems}$ & Stem split time (years) & 411700 & 381400--410700 \\
    $T_{stem1}$ & Stem 1 split time into stem 1E and stem 1S (years) & 355400 & 154300--320400 \\
    $m_{Nama-MSL}$ & Nama--Mende symmetric migration rate & $0.361\times10^{-5}$ & $0.205\times10^{-5}$--$0.697\times10^{-5}$ \\
    $m_{Nama-EA}$ & Nama--East Africa symmetric migration rate & $4.04\times10^{-5}$ & $3.90\times10^{-5}$--$4.42\times10^{-5}$ \\
    $m_{MSL-EA}$ & Mende--East Africa migration rate & $19.3\times10^{-5}$ & $19.8\times10^{-5}$--$20.8\times10^{-5}$ \\
    $m_{EA-GBR}$ & East Africa--Europe migration rate & $3.76\times10^{-5}$ & $3.34\times10^{-5}$--$3.95\times10^{-5}$ \\
    $m_{EA-EA}$ & Intra-East Africa migration rate & $37.2\times10^{-5}$ & $35.2\times10^{-5}$--$39.6\times10^{-5}$ \\
    $f_{GBR \rightarrow EP}$ & Ancestry proportion of East African agriculturalists from GBR 12ka ($1-f$ from Gumuz) & 0.648 & 0.641--0.661 \\
    $f_{EP \rightarrow Nama}$ & Ancestry proportion from EA pastoralists to Nama 2ka & 0.258 & 0.245--0.264 \\
    $f_{GBR \rightarrow Nama}$ & Ancestry proportion from Europeans to Nama 10 generations ago & 0.156 & 0.152--0.161 \\
    $m_{stems}$ & Stem 1--stem 2 migration rate & $0$ & fixed \\
    $T_{Nama}$ & Time of Nama merger event & 115400 & 101400--116800 \\
    $f_{stem 2 \rightarrow Nama}$ & Proportion of stem 2 ancestry making up initial Nama lineage ($1-f$ from stem 1S) & 0.693 & 0.706--0.769 \\
    $T_{EA}$ & Time of East Africa merger event & 92940 & 94500--10140 \\
    $f_{stem 2 \rightarrow EA}$ & Proportion of stem 2 ancestry making up initial East Africa lineage ($1-f$ from stem 1E) & 0.502 & 0.510--0.564 \\
    $T_{MSL}$ & Time of secondary admixture from stem 2 to Mende & 23490 & 22920--25510 \\
    $f_{stem 2 \rightarrow MSL}$ & Proportion of ancestry from secondary stem 2 admixture to Mende & 0.170 & 0.174--0.196 \\
    \bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{
    \label{tab:supp-merger-with-stem-migration}
    \textbf{Best-fit parameters from the Merger-With-Stem-Migration model.}
    Inferred values are scaled to physical units assuming a generation time of
    29 years. This model gave a log-likelihood of -102,633.
    \aprcomment{check these at end for each model}
    \aprcomment{if the slightly shift, also fix elsewhere in supp and main text}
}
\centering
\begin{tabular}[t]{rP{8cm}cc}
    \toprule
    Parameter & Description & Value & Std. err.\\
    \midrule
    $N_e$ & Ancestral effective population size & 10960 & 64500--11910 \\
    $N_{stem1}$ & Size of stem 1 lineage between Neanderthal split and stem 1E--stem 1S split & 100 & 100--851 \\
    $N_{stem2}$ & Size of stem 2 lineage & 24100 & 22130--26340 \\
    $N_{Nama_0}$ & Initial Nama size & 13530 & 11500--13560 \\
    $N_{Nama_F}$ & Final Nama size & 219 & 214--231 \\
    $N_{MSL_0}$ & Initial Mende size & 11500 & 9690--11290 \\
    $N_{MSL_F}$ & Final Mende size & 27390 & 26820--32280 \\
    $N_{EA}$ & Size of East African Branch & 9150 & 7970--8860 \\
    $N_{Gumuz_F}$ & Final Gumuz size & 3390 & 3270--3570 \\
    $N_{EP}$ & East African agriculturist size & 13500 & 12780--13500 \\
    $N_{GBR_0}$ & Initial British size & 950 & 928--1002 \\
    $N_{GBR_F}$ & Final British size & 12370 & 11850--12600 \\
    $N_{Neand}$ & Neanderthal size & 2510 & 2100--2730 \\
    $T_{stems}$ & Stem split time (years) & 1475000 & 695000--1785000 \\
    $T_{stem1}$ & Stem 1S--stem 1E split time (years) & 479100 & 275600--478300 \\
    $m_{Nama-MSL}$ & Nama--Mende symmetric migration rate & $1.06\times10^{-5}$ & $0.66\times10^{-5}$--$1.57\times10^{-5}$ \\
    $m_{Nama-EA}$ & Nama--East Africa symmetric migration rate & $4.49\times10^{-5}$ & $4.45\times10^{-5}$--$5.52\times10^{-5}$ \\
    $m_{MSL-EA}$ & Mende--East Africa migration rate & $19.6\times10^{-5}$ & $20.0\times10^{-5}$--$21.7\times10^{-5}$ \\
    $m_{EA-GBR}$ & East Africa--Europe migration rate & $3.85\times10^{-5}$ & $3.25\times10^{-5}$--$4.27\times10^{-5}$ \\
    $m_{EA-EA}$ & Intra-East Africa migration rate & $36.0\times10^{-5}$ & $33.7\times10^{-5}$--$40.9\times10^{-5}$ \\
    $f_{GBR \rightarrow EP}$ & Ancestry proportion of East African agriculturalists from GBR 12ka ($1-f$ from Gumuz) & 0.639 & 0.636--0.667 \\
    $f_{EP \rightarrow Nama}$ & Ancestry proportion from EA pastoralists to Nama 2ka & 0.259 & 0.237--0.267 \\
    $f_{GBR \rightarrow Nama}$ & Ancestry proportion from Europeans to Nama 10 generations ago & 0.159 & 0.154--0.167 \\
    $m_{stems}$ & Stem 1--stem 2 migration rate & $11.8\times10^{-5}$ & $3.95\times10^{-5}$--$12.3\times10^{-5}$ \\
    $T_{Nama}$ & Time of Nama merger event & 121600 & 101300--124600 \\
    $f_{stem 2 \rightarrow Nama}$ & Proportion of stem 2 ancestry making up initial Nama lineage ($1-f$ from stem 1S) & 0.692 & 0.694--0.758 \\
    $T_{EA}$ & Time of East Africa merger event & 96500 & 95400--111000 \\
    $f_{stem 2 \rightarrow EA}$ & Proportion of stem 2 ancestry making up initial East Africa lineage ($1-f$ from stem 1E) & 0.511 & 0.514--0.589 \\
    $T_{MSL}$ & Time of secondary admixture from stem 2 to Mende & 24890 & 22210--26300 \\
    $f_{stem 2 \rightarrow MSL}$ & Proportion of ancestry from secondary stem 2 admixture to Mende & 0.182 & 0.182--0.211 \\
    \bottomrule
\end{tabular}
\end{table}
